% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- introduction file header -----------------------


%------------------------------------------------------------------------- 


% the code below specifies where the figures are stored
%\ifpdf
%    \graphicspath{{Content/figures/PNG/}{Content/figures/JPG/}{Content/figures/PDF/}{Content/figures/}}
%\else
%    \graphicspath{{Content/figures/EPS/}{Content/figures/}}
%\fi
\graphicspath{{ResearchResults/figures/}{ResearchResults/figures/JPG/}{ResearchResults/figures/PDF/}{ResearchResults/figures/}{ResearchResults/figures/EPS/}}

\chapter{Interoperable Architecture for Collaborative Augmented Reality}
\chaptermark{Architecture}
\label{chap:arch}

In this Chapter the \arch{} architecture is presented. \arch{} is an interoperable architecture that simplifies the creation of collaborative AR applications, enables multi-user functionalities and provides advanced mechanisms for data analysis and visualisation. The architecture aims to answer the research gaps identified in Chapter \ref{chap:sota} as well as the design objectives identified analysing the answers to a survey completed by 47 primary and secondary school teachers. The architecture has been validated through the development of 3 proofs of concept. \arch{} provides a more mature technological ecosystem that fosters the emergence of AR applications for education and their inclusion in existing school programs.

\input{ResearchResults/architecture}

\chapter{Validation of the Proposed Architecture}
\chaptermark{Validation}
\label{chap:eval}

This Chapter presents an application called \appname{}, a multiplatform AR application for education. It is based on the \arch{} architecture presented in Chapter \ref{chap:arch}, and designed with the help of secondary school teachers. The app provides interoperability, multi-user support, integration with LMSs and data analytics capabilities, thus simplifying the development of collaborative AR learning experiences. \appname{} was developed with the purpose of validating, in an educational environment, the \arch{} architecture.
The application has been tested by \numstudents/ students and 3 teachers from \numschools/ different educational institutions to evaluate the usability as well as the impact of collaboration functionalities in the engagement of the students.

\input{ResearchResults/evaluation}
